Recursive algorithms and branching for nonmaximal embeddings

### 2010-11-18 AffineLieAlgebras/DualFunctions/article.tex
I've started to write the paper from the MQFT2010 talk.

The paper is organised as follows. In section \ref{sec:finite-dimens-lie} we describe the recurrent process that gives rise to the systems of the orthogonal polynomials. This construction in some sense is not new (e.g. see \cite{2010arXiv1001.3683N}), but we explicitly describe the dual system of functions.  The section \ref{sec:affine-lie-algebras} is devoted to the systems of special functions connected with the affine Lie algebras. We establish the modular properties of the dual functions. Possible applications in physics are described in the next section \ref{sec:poss-appl}. We discuss the importance of the tensor product decompositions for the integrable systems which follows from the Hecke duality. Also we consider conformal field theory.
In the conclusion \ref{sec:conclusion} we overview the results and state some open questions.

@@

\section{Finite-dimensional Lie algebras}
\label{sec:finite-dimens-lie}
Consider the irreducible highest-weight module $L^{(\mu)}_{\mathfrak{g}}$ of the simple Lie algebra $\mathfrak{g}$. 
The recurrent relation for the weight multiplicities can be written from the Weyl character formula \cite{lyakhovsky1996rra}:
\begin{equation}
  \label{eq:1}
  \mathrm{ch} L^{(\mu)}_{\mathfrak{g}}=\frac{\sum_{w\in W}\epsilon(w) e^{w(\mu+\rho)-\rho}}{\sum_{w\in W}\epsilon(w) e^{w \rho -\rho}}
\end{equation}
Let us limit our consideration to the fundamental domain (main Weyl chamber). The left side of this equation can be written as
\begin{equation}
  \label{eq:2}
   \mathrm{ch} L^{(\mu)}_{\mathfrak{g}}=\sum_{\lambda\in \bar{C}_{\mathfrak{g}}} m^{(\mu)}_{\lambda} e^{\lambda}
\end{equation}
Multiplying the equation \eqref{eq:1} by the denominator  and collecting the common terms we get the recurrent relation for the weight multiplicities $m^{(\mu)}_{\lambda}$:
\begin{equation}
  \label{eq:3}
  \sum_{w\in W} \epsilon(w) m^{(\mu)}_{\lambda-w\rho+\rho}=\sum_{w\in W}\epsilon(w) \delta_{\lambda,w(\mu+\rho)-\rho}
\end{equation}
If we consider the weights inside the main Weyl chamber the sum on the right-hand side is reduced to the single term $\delta_{\lambda,\mu}$. For the weight $\lambda-w\rho+\rho\not\in \bar{C}_{\mathfrak{g}}$ there exists weight $\xi\in \bar{C}_{\mathfrak{g}}$ and Weyl group element $v\in W$ such that $v\xi=\lambda-w\rho+\rho$, so we can limit the computation of weight multiplicities by the main Weyl chamber. This procedure was called ``folding of the fan'' in \cite{il2010folded}.
Now we want to rewrite the equation \eqref{eq:3} in matrix form similarly to \cite{2010arXiv1001.3683N}.
We need to choose the ordering on the set of weights in the closure of fundamental domain $\bar{C}_{\mathfrak{g}}$. We use the following natural procedure. Consider some ordering of the simple roots (e.g. the numbering of nodes of Dynkin diagram from left to right). Let's denote the weights by the Dynkin labels. Then the task is to enumerate the set $\{0\}\cup Z_{+}^r$. $[0,\dots,0]$ get the index 1, $[0,\dots,0,1]$ the index 2, $[0,\dots,0,1,0]$ - 3 etc.
\section{Affine Lie algebras}
\label{sec:affine-lie-algebras}
Consider the (affine) Lie algebra $\mathfrak{g}$  and its reductive subalgebra $\af\subset \mathfrak{g}$. Introduce the following objects.
\begin{definition}
  Denote by $\Delta^{+}_{\bot}=\left\{\alpha\in \Delta^{+}_{\mathfrak{g}}:\forall \beta\in\Delta_{\af} \; \alpha\bot \beta\right\}$ the subset of the positive roots of $\mathfrak{g}$ orthogonal to the root system of $\af$.
\end{definition}

>>>>

The paper is organised as follows. In section \ref{sec:finite-dimens-lie} we describe the recurrent process that gives rise to the systems of the orthogonal polynomials. This construction in some sense is not new (e.g. see \cite{2010arXiv1001}), but we explicitly describe the dual system of functions.  The section \ref{sec:affine-lie-algebras} is devoted to the systems of special functions connected with the affine Lie algebras. We establish the modular properties of the dual functions. Possible applications in physics are described in the next section \ref{sec:poss-appl}. We discuss the importance of the tensor product decompositions for the integrable systems which follows from the Hecke duality. Also we consider conformal field theory.
In the conclusion \ref{sec:conclusion} we overview the results and state some open questions.

@@

\section{Finite-dimensional Lie algebras}
\label{sec:finite-dimens-lie}
Consider the irreducible highest-weight module $L^{(\mu)}_{\mathfrak{g}}$ of the simple Lie algebra $\mathfrak{g}$. 
The recurrent relation for the weight multiplicities can be written from the Weyl character formula \cite{lyakhovsky1996rra}:
\begin{equation}
  \label{eq:1}
  \mathrm{ch} L^{(\mu)}_{\mathfrak{g}}=\frac{\sum_{w\in W}\epsilon(w) e^{w(\mu+\rho)-\rho}}{\sum_{w\in W}\epsilon(w) e^{w \rho -\rho}}
\end{equation}
Let us limit our consideration to the fundamental domain (main Weyl chamber). The left side of this equation can be written as
\begin{equation}
  \label{eq:2}
   \mathrm{ch} L^{(\mu)}_{\mathfrak{g}}=\sum_{\lambda\in \bar{C}_{\mathfrak{g}}} m^{(\mu)}_{\lambda} e^{\lambda}
\end{equation}
Multiplying the equation \eqref{eq:1} by the denominator  and collecting the common terms we get the recurrent relation for the weight multiplicities $m^{(\mu)}_{\lambda}$:
\begin{equation}
  \label{eq:3}
  \sum_{w\in W} \epsilon(w) m^{(\mu)}_{\lambda-w\rho+\rho}=\sum_{w\in W}\epsilon(w) \delta_{\lambda,w(\mu+\rho)-\rho}
\end{equation}
If we consider the weights inside the main Weyl chamber the sum on the right-hand side is reduced to the single term $\delta_{\lambda,\mu}$. For the weight $\lambda-w\rho+\rho\not\in \bar{C}_{\mathfrak{g}}$ there exists weight $\xi\in \bar{C}_{\mathfrak{g}}$ and Weyl group element $v\in W$ such that $v\xi=\lambda-w\rho+\rho$, so we can limit the computation of weight multiplicities by the main Weyl chamber. This procedure was called ``folding of the fan'' in \cite{il2010folded}.
Now we want to rewrite the equation \eqref{eq:3} in matrix form similarly to \cite{2010arXiv1001}.
We need to choose the ordering on the set of weights in the closure of fundamental domain $\bar{C}_{\mathfrak{g}}$. We use the following natural procedure. Consider some ordering of the simple roots (e.g. the numbering of nodes of Dynkin diagram from left to right). Let's denote the weights by the Dynkin labels. Then the task is to enumerate the set $\{0\}\cup Z_{+}^r$. $[0,\dots,0]$ get the index 1, $[0,\dots,0,1]$ the index 2, $[0,\dots,0,1,0]$ - 3 etc.
\section{Affine Lie algebras}
\label{sec:affine-lie-algebras}
Consider the (affine) Lie algebra $\mathfrak{g}$  and its reductive subalgebra $\af\subset \mathfrak{g}$. Introduce the following objects.
\begin{definition}
  Denote by $\Delta^{+}_{\bot}=\left\{\alpha\in \Delta^{+}_{\mathfrak{g}}:\forall \beta\in\Delta_{\af} \; \alpha\bot \beta\right\}$ the subset of the positive roots of $\mathfrak{g}$ orthogonal to the root system of $\af$.
\end{definition}

### 2010-11-18 AffineLieAlgebras/InformalNotes/article.tex
I've started to write the paper from the MQFT2010 talk.

\documentclass[12pt]{iopart}
%\usepackage{ucs}
%\usepackage{amsmath,amssymb,amsthm}
\usepackage{iopams,amssymb,amsthm}
%\usepackage{pb-diagram}
\usepackage{multicol}
%\usepackage[utf8x]{inputenc}
%\usepackage[russian]{babel}
%\usepackage{cmap}
\usepackage{color}
\usepackage{graphicx}
%\usepackage{epstopdf}
%\usepackage{graphicx}
%\pagestyle{plain}
%\usepackage[unicode,verbose,plainpages=false]{hyperref}
\usepackage{hyperref}

@@

Usually only the maximal reductive subalgebras $\frak{a} \subset \frak{g} $ are considered since the case of non-maximal subalgebra
can be obtained using the chain of maximal injections. In this paper we find out that the recurrent properties
for branching coefficients can be explicitly formulated for an arbitrary reductive subalgebra.
The principal point is to consider the subalgebra $\af$ together with its
counterpart $\afb$ "orthogonal" to $\af$ with respect to the Killing form.
For any reductive $\af$ the subalgebra $\afb \subset \frak{g} $ is a regular and reductive.
For a highest weight module $L^{\left( \mu \right)}$ and orthogonal pair of subalgebras
$\left(  \af, \afb \right)$ we consider
%In the the group algebra $\mathcal{E}\left( \frak{g} \right)$ we consider
the so called singular element $\Psi^{\left( \mu \right)}$ (the numerator
in the Weyl character formula
$ch\left( L^{\mu }\right) =\frac{\Psi ^{\left( \mu \right) }}{\Psi ^{\left( 0\right) }}$, see for example \cite{humphreys1978rth})
the Weyl denominator $\Psi ^{\left( 0\right) }_{\afb}$ and the projection
$\Psi ^{\left( \mu \right) }_{\left(  \af, \afb \right)}=\pi_{\af}\frac{\Psi ^{\left( \mu \right) }_{\frak{g}}}{\Psi ^{\left( 0\right) }_{\afb}}$.
We prove that for any highest weight module $L^{\left( \mu \right)}$ and orthogonal pair of subalgebras
$\left(  \af, \afb \right)$ the element $\Psi ^{\left( \mu \right) }_{\left(  \af, \afb \right)}$ has a decomposition with respect to
the set of Weyl numerators of $\afb$.
This decomposition provides the possibility to construct the recurrent property for the branching coefficients corresponding
to the injection $\frak{a} \longrightarrow \frak{g} $.
This property is formulated in
terms of a specific element of the group algebra $\mathcal{E}\left( \frak{g} \right)$ called "the subtracted injection fan".
Using this tool we formulate a simple and
explicit algorithm for branching coefficients computations applicable for an arbitrary (maximal or non-maximal)
subalgebras of finite-dimensional and affine Lie algebras.
In the case of maximal embedding the corresponding fan becomes unsubtracted and the recurrent relations described
earlier (see the paper \cite{ilyin812pbc}
and the references therein) are reobtained.

@@

In the Weyl-Kac formula both numerator and denominator  can be considered
as formal elements containing the singular weights of the Verma modules $V^{\xi}$
with the highest weights $\xi=\mu$ and $\xi=0$ \cite{humphreys1978rth}.
We attribute singular numerators to the corresponding integrable modules $L^{\mu }$
and $L_{\af}^{\nu }$:
\begin{equation*}
\Psi ^{\left( \mu \right) }:=\sum\limits_{w\in W}\epsilon (w)e^{w\circ (\mu +\rho )-\rho },
\end{equation*}
\begin{equation*}
\Psi _{ \af}^{\left( \nu \right) }:=
\sum\limits_{w\in W_{\af}}\epsilon (w)e^{w\circ (\nu +\rho
_{_{\af}})-\rho _{_{\af}}}.
\end{equation*}
and use the Weyl-Kac formula in the form
\begin{equation}
\label{Weyl-Kac2}
ch\left( L^{\mu }\right) =\frac{\Psi ^{\left( \mu \right) }}
{\Psi ^{\left( 0 \right) }}=\frac{\Psi ^{\left( \mu \right) }}{R}.
\end{equation}

@@

Consider the right-hand side of relation (\ref{eq:4}).
The numerator there describes the branching in terms of singular elements and
it is reasonable to expand it as an element of $\mathcal{E}\left( \frak{g} \right)$:
\begin{equation}
  \label{eq:21}
  \sum_{\nu \in \bar{C_{\mathfrak{a}}}}b_{\nu }^{\left( \mu \right) }\Psi _{\left( \frak{%
        a}\right) }^{\left( \nu \right) }=\sum_{\lambda \in P_{\af}}k_{\lambda
  }^{\left( \mu \right) }e^{\lambda }.
\end{equation}
Here the coefficients $k_{\lambda}^{\left( \mu \right) }$ are integer and their signes
depend on the length (see \cite{humphreys1978rth})  of the Weyl group elements in
$\Psi _{\left( \frak{a}\right) }^{\left( \nu \right) }$. The important property of
$k_{\lambda}^{\left( \mu \right) }$'s is that they coinside with the branching coefficients
for all weights $\nu$ inside the main Weil chamber:
\begin{equation}
%  \label{eq:20}
  b^{(\mu)}_{\nu}=k^{(\mu)}_{\nu} \; \mbox{for} \; \nu\in \bar{C}_{\mathfrak{a}}.
\label{eq:21-1}
\end{equation}
We call the coefficients $k_{\lambda}$ --- the anomalous branching coefficients
(see also \cite{ilyin812pbc}).

>>>>

\documentclass[12pt]{iopart}
%\usepackage{ucs}
%\usepackage{amsmath,amssymb,amsthm}
\usepackage{iopams,amssymb,amsthm}
%\usepackage{pb-diagram}
\usepackage{multicol}
%\usepackage[utf8x]{inputenc}
%\usepackage[russian]{babel}
%\usepackage{cmap}
\usepackage{color}
\usepackage{graphicx}
\usepackage{epstopdf}
%\usepackage{graphicx}
%\pagestyle{plain}
%\usepackage[unicode,verbose,plainpages=false]{hyperref}
\usepackage{hyperref}

@@

Usually only the maximal reductive subalgebras $\frak{a} \subset \frak{g} $ are considered since the case of non-maximal subalgebra
can be obtained using the chain of maximal injections. In this paper we find out that the recurrent properties
for branching coefficients can be explicitly formulated for an arbitrary reductive subalgebra.
The principal point is to consider the subalgebra $\af$ together with its
counterpart $\afb$ "orthogonal" to $\af$ with respect to the Killing form.
For any reductive $\af$ the subalgebra $\afb \subset \frak{g} $ is a regular and reductive.
For a highest weight module $L^{\left( \mu \right)}$ and orthogonal pair of subalgebras
$\left(  \af, \afb \right)$ we consider
%In the the group algebra $\mathcal{E}\left( \frak{g} \right)$ we consider
the so called singular element $\Psi^{\left( \mu \right)}$ (the numerator
in the Weyl character formula
$ch\left( L^{\mu }\right) =\frac{\Psi ^{\left( \mu \right) }}{\Psi ^{\left( 0\right) }}$, see for example \cite{humphreys1997introduction})
the Weyl denominator $\Psi ^{\left( 0\right) }_{\afb}$ and the projection
$\Psi ^{\left( \mu \right) }_{\left(  \af, \afb \right)}=\pi_{\af}\frac{\Psi ^{\left( \mu \right) }_{\frak{g}}}{\Psi ^{\left( 0\right) }_{\afb}}$.
We prove that for any highest weight module $L^{\left( \mu \right)}$ and orthogonal pair of subalgebras
$\left(  \af, \afb \right)$ the element $\Psi ^{\left( \mu \right) }_{\left(  \af, \afb \right)}$ has a decomposition with respect to
the set of Weyl numerators of $\afb$.
This decomposition provides the possibility to construct the recurrent property for the branching coefficients corresponding
to the injection $\frak{a} \longrightarrow \frak{g} $.
This property is formulated in
terms of a specific element of the group algebra $\mathcal{E}\left( \frak{g} \right)$ called "the subtracted injection fan".
Using this tool we formulate a simple and
explicit algorithm for branching coefficients computations applicable for an arbitrary (maximal or non-maximal)
subalgebras of finite-dimensional and affine Lie algebras.
In the case of maximal embedding the corresponding fan becomes unsubtracted and the recurrent relations described
earlier (see the paper \cite{ilyin812pbc}
and the references therein) are reobtained.

@@

In the Weyl-Kac formula both numerator and denominator  can be considered
as formal elements containing the singular weights of the Verma modules $V^{\xi}$
with the highest weights $\xi=\mu$ and $\xi=0$ \cite{humphreys1997introduction}.
We attribute singular numerators to the corresponding integrable modules $L^{\mu }$
and $L_{\af}^{\nu }$:
\begin{equation*}
\Psi ^{\left( \mu \right) }:=\sum\limits_{w\in W}\epsilon (w)e^{w\circ (\mu +\rho )-\rho },
\end{equation*}
\begin{equation*}
\Psi _{ \af}^{\left( \nu \right) }:=
\sum\limits_{w\in W_{\af}}\epsilon (w)e^{w\circ (\nu +\rho
_{_{\af}})-\rho _{_{\af}}}.
\end{equation*}
and use the Weyl-Kac formula in the form
\begin{equation}
\label{Weyl-Kac2}
ch\left( L^{\mu }\right) =\frac{\Psi ^{\left( \mu \right) }}
{\Psi ^{\left( 0 \right) }}=\frac{\Psi ^{\left( \mu \right) }}{R}.
\end{equation}

@@

Consider the right-hand side of relation (\ref{eq:4}).
The numerator there describes the branching in terms of singular elements and
it is reasonable to expand it as an element of $\mathcal{E}\left( \frak{g} \right)$:
\begin{equation}
  \label{eq:21}
  \sum_{\nu \in \bar{C_{\mathfrak{a}}}}b_{\nu }^{\left( \mu \right) }\Psi _{\left( \frak{%
        a}\right) }^{\left( \nu \right) }=\sum_{\lambda \in P_{\af}}k_{\lambda
  }^{\left( \mu \right) }e^{\lambda }.
\end{equation}
Here the coefficients $k_{\lambda}^{\left( \mu \right) }$ are integer and their signes
depend on the length (see \cite{humphreys1997introduction})  of the Weyl group elements in
$\Psi _{\left( \frak{a}\right) }^{\left( \nu \right) }$. The important property of
$k_{\lambda}^{\left( \mu \right) }$'s is that they coinside with the branching coefficients
for all weights $\nu$ inside the main Weil chamber:
\begin{equation}
%  \label{eq:20}
  b^{(\mu)}_{\nu}=k^{(\mu)}_{\nu} \; \mbox{for} \; \nu\in \bar{C}_{\mathfrak{a}}.
\label{eq:21-1}
\end{equation}
We call the coefficients $k_{\lambda}$ --- the anomalous branching coefficients
(see also \cite{ilyin812pbc}).
